Matching Bullets

### 2015-12-16 imaging-paper.Rnw
update to elsevier format

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Large Colored Title Article
% LaTeX Template
% Version 1.1 (25/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[DIV=calc, paper=letter, fontsize=10pt, twocolumn]{scrartcl}	 % A4 paper and 11pt font size

\usepackage[USenglish]{babel}
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{float}
%\usepackage{afterpage}

@@

% Headers - all currently empty
\lhead{}
\chead{}
\rhead{}

% Footers
\lfoot{}
\cfoot{}
\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % "Page 1 of 2"

\renewcommand{\headrulewidth}{0.0pt} % No header rule
\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule

@@

\usepackage{lettrine} % Package to accentuate the first letter of the text
\newcommand{\initial}[1]{ % Defines the command and style for the first letter
\lettrine[lines=3,lhang=0.3,nindent=0em]{
\color{DarkGoldenrod}
{\textsf{#1}}}{}}

\usepackage{subcaption}

@@

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\usepackage{titling} % Allows custom title configuration

\newcommand{\HorRule}{\color{DarkGoldenrod} \rule{\linewidth}{1pt}} % Defines the gold horizontal rule around the title

\pretitle{\vspace{-30pt} \begin{flushleft} \HorRule \fontsize{50}{50} \usefont{OT1}{phv}{b}{n} \color{DarkRed} \selectfont} % Horizontal rule before the title

\title{Matching Bullets} % Your article title

\posttitle{\par\end{flushleft}\vskip 0.5em} % Whitespace under the title

\preauthor{\begin{flushleft}\large \lineskip 0.5em \usefont{OT1}{phv}{b}{sl} \color{DarkRed}} % Author font configuration

\author{Eric Hare, Heike Hofmann} % Your name

\postauthor{\footnotesize \usefont{OT1}{phv}{m}{sl} \color{Black} % Configuration for the institution name

Department of Statistics and Statistical Laboratory,
Iowa State University% Your institution

\par\end{flushleft}\HorRule} % Horizontal rule after the title

\date{} % Add a date here if you would like one to appear underneath the title block

@@

\maketitle % Print the title

\thispagestyle{fancy} % Enabling the custom headers/footers for the first page

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

% The first character should be within \initial{}
\initial{F}\textbf{orensic examiners try to determine whether two bullets were fired from the same barrel by analyzing them under a microscope in controlled conditions. Can we make the task simpler?}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\paragraph{Keywords:} Data visualization, Statistical graphics, Statistical computing.

@@

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residuala}Residual structure of a surface crosscut at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residualb} Residual structure of a surface crosscut at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed crosscut ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Crosscuts at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Adjusted (bullet 1-5 horizontally shifted) crosscuts at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface crosscuts at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

\subsection{Horizontal Alignment}
Using (for now) a single crosscut across the surface, we extract the values and remove the class characteristics using a loess fit as before. This enables us to focus on the individual characterists. Striation marks show up in this representations as peaks and valleys. Figure~\ref{fig:cross100} shows lines of the surface crosscut at $x = 100$. A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.
This horizontal shift is based on the cross-correlation between the two surface crosscuts. Let $f(t)$ and $g(t)$ define the values of the crosscut at $t$, then the cross-correlation between $f$ and $g$ at lag $k$ is defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

\begin{figure}[hbtp]
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two crosscuts shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the crosscuts. Figure~\ref{fig:crosscutX} shows the lag-shifted crosscuts.}
\end{figure}

\subsection{Varying crosscuts}
The crosscut at which a comparison is done matters -- as crosscuts are further apart, more pronounced differences between the crosscuts show up. This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which crosscut level was used in a matching, but on the other hand we can use multiple crosscuts of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of crosscuts for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$. These are compared to a crosscut at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two crosscuts. However, the match quickly deteriorates with increasing distance to non-matches at a crosscut of $x = 300\mu m$ or higher.  Only if the crosscut is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same crosscut levels due to manual alignments in microscopes, we should take crosscut levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 450, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
             lop[[7]],
             ncol = 1)
@
\caption{\label{fig:crosscuts}Overview of the variations in the surface along different crosscuts. Crosscut at $x = 100$ is compared to crosscuts every 50$\mu m$. With every step away from the original crosscut site, the number of differences between the crosscuts increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher. }
\end{figure}
%\afterpage{\clearpage}

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
             lop[[4]], lop[[5]], 
             ncol=1)
@
\caption{\label{fig:crosscuts2}Varying crosscuts for barrel 3, bullet 1. Initially, the match between crosscuts 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet. At a crosscut level of $175\mu m$ the bullet pattern stabilizes. For this land, matches should not be attempted at lower crosscuts.}
\end{figure}
An opportunity that arises from comparing multiple crosscuts of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the crosscut is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well. In the approach here, we keep increasing the crosscut level until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3.
`Stability' is here defined as two aligned crosscuts chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9.

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Using a rolling median peaks and grooves are identified for each crosscut. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bullet~.) +
    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Rectangles in the back identify a line on one of the bullets.  Matching striation marks are indicated by color filled rectangles and arked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Crosscut values (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between crosscuts (c).}
\end{figure}
%\afterpage{\clearpage}

@@

Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet and pick the lowest crosscut level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet. 
%\hh{(assumption here: lower crosscut levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}

>>>>

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of Forensic Science International}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[USenglish]{babel}
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document
\usepackage{float}
\usepackage{subcaption}

@@

@@

@@

@@

\begin{frontmatter}

\title{Matching Bullets}
\tnotetext[mytitlenote]{Resources for this manuscript, including R code to reproduce any of the models, tables and figures, can be found at  \url{https://github.com/erichare/imaging-paper}.}

%% Group authors per affiliation:
\author{Eric Hare, Heike Hofmann, Alicia Carriquiry}
\address{Department of Statistics and Statistical Laboratory, Iowa State University}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Eric Hare}
\ead{erichare@iastate.edu}

\address[mymainaddress]{Department of Statistics and Statistical Laboratory, Iowa State University}
\address[mysecondaryaddress]{secondary address -- CS?}

\begin{abstract}
In 2009, the National Academy of Sciences published a report questioning the scientific validity of many forensic methods including firearm examination. Firearm examination is a forensic tool used to determine whether two bullets were fired from the same gun barrel. During the firing process, rifling, manufacturing defects, and impurities in the barrel create striation marks on the bullet. Identifying these striation markings in an attempt to match two bullets is the primary goal. We propose a framework for the analysis of the 3D surface measurements which transcribes the markings into a 2D plotting framework. This makes identification of matches easier and more rooted in science for both the expert and the lay person. The automatic matching routine we propose manages to correctly predict 132 out of 180 known bullet land to bullet land matches.
\end{abstract}

\begin{keyword}
\hh{XXX not the right keywords yet} Data visualization, Statistical graphics, Statistical computing.
\end{keyword}

\end{frontmatter}

\linenumbers

@@

The remainder of this paper is structured as follows: We first discuss two methods of modeling the class structure of the bullet surfaces. We then describe a web application for visualizing and working with bullet surfaces. We proceed to describing an automatic matching routine which we evaluate on a small database of reference bullets.

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residuala}Residual structure of a surface cross section at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:residualb} Residual structure of a surface cross section at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\linewidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

@@

\begin{figure}[hbtp]
\begin{subfigure}[t]{\linewidth}
\caption{\label{fig:loess_step1}Step 1 the loess fit algorithm. For a fixed cross section ($x$ held constant at 243.75 $\mu m$) of surface measurements for bullet 1-5 are plotted across different values of $y$.}
<<loess_step1, echo=FALSE, fig.width=10, fig.height=3.5, out.width='\\linewidth', warning=FALSE,  message=FALSE>>=
br111_bullet <- get_bullet(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
br121_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-1.x3p", sep = "/"))
br122_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-2.x3p", sep = "/"))
br123_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-3.x3p", sep = "/"))
br124_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-4.x3p", sep = "/"))
br125_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-5.x3p", sep = "/"))
br126_bullet <- get_bullet(paste(datadir,"Br1 Bullet 2-6.x3p", sep = "/"))

@@

lines <- matches %>% group_by(lineid) %>% summarise(
  meany = mean(y, na.rm=T),
  miny = min(y, na.rm=T),
  maxy = max(y, na.rm=T) + 1.5625,
  match = isMatch(lineid, type),
  type = type1[1]
)
lines <- subset(lines, lineid != 0)
@
\begin{figure}[hbtp]
\centering
\begin{subfigure}[t]{\linewidth}
\caption{Cross section at $x = 100$.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(.9, .2),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Adjusted (bullet 1-5 horizontally shifted) cross sections at $x = 100$.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
# qplot(y, resid, group=bullet, colour=bullet, data=lofX, 
#       geom="line", alpha=I(0.6)) + 
#   scale_colour_brewer(palette="Set1") + theme_bw() +
#   theme(legend.position="none")
qplot(y, resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position="none") + ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Lines showing surface cross sections at $x = 100$. A horizontal shift of the values of bullet 1-5 to the right shows the similarity of the striation marks.}
\end{figure}

\subsection{Horizontal Alignment}
Using (for now) a single cross section across the surface, we extract the values and remove the class characteristics using a loess fit as before. This enables us to focus on the individual characterists. Striation marks show up in this representations as peaks and valleys. Figure~\ref{fig:cross100} shows lines of the surface cross section at $x = 100$. A horizontal shift of one of the lines (result shown in (b)) emphasizes the strong similarities.
This horizontal shift is based on the cross-correlation between the two surface cross sections. Let $f(t)$ and $g(t)$ define the values of the cross section at $t$, then the cross-correlation between $f$ and $g$ at lag $k$ is defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

\begin{figure}[hbtp]
<<ccf, echo=FALSE, fig.width = 7, fig.height = 3, out.width = '\\linewidth'>>=
lag <- ccf$lag[which.max(ccf$acf)]
qplot(x=ccf$lag, xend=ccf$lag, y = 0, yend= ccf$acf, geom="segment", colour = I("grey50")) + 
  theme_bw() + ylab("Correlation") + xlab("Lag k") +
  geom_segment(aes(x = lag, xend = lag, y = 0, yend =  max(ccf$acf)), colour="black") + 
  scale_x_continuous(breaks = c(-100, -50, lag, 0, 50, 100),
                     minor_breaks = c(-75,-25,25, 75))
@
\caption{\label{fig:ccf}Cross-correlation between the two cross sections shown in Figure~\ref{fig:crosscut} at lags between -100 and 100. At a lag of -10 the correlation peaks, indicating the largest amount of agreement between the cross sections. Figure~\ref{fig:crosscutX} shows the lag-shifted cross sections.}
\end{figure}

\subsection{Varying cross sections}
The cross section at which a comparison is done matters -- as cross sections are further apart, more pronounced differences between the cross sections show up. This poses both a caveat to matching attempts as well as an opportunity: on the one hand we have to be aware of which cross section level was used in a matching, but on the other hand we can use multiple cross sections of the same bullet for an initial assessment of its quality. Both of these situations are assessed and discussed in more detail below: 
Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 450$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
<<crosscuts-vary, echo=FALSE, fig.width = 7, fig.height = 14, out.width = '\\linewidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 450, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]], 
             lop[[7]],
             ncol = 1)
@
\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections. Cross section at $x = 100$ is compared to cross sections every 50$\mu m$. With every step away from the original cross section site, the number of differences between the cross sections increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher. }
\end{figure}
%\afterpage{\clearpage}

@@

grid.arrange(lop[[1]], lop[[2]], lop[[3]], 
             lop[[4]], lop[[5]], 
             ncol=1)
@
\caption{\label{fig:crosscuts2}Varying cross sections for barrel 3, bullet 1. Initially, the match between cross sections 25$\mu m$ apart is affected strongly by break off at the bottom of the bullet. At a cross section level of $175\mu m$ the bullet pattern stabilizes. For this land, matches should not be attempted at lower cross sections.}
\end{figure}
An opportunity that arises from comparing multiple cross sections of a bullet's surface measurements is the following: by comparing cross sections that are not too far apart -- maybe 25 or 50 $\mu m$ for example -- we get an indication whether the cross section is in a rapidly changing section of the surface, indicative of a break-off, or in a stable section, where we can expect to find matches to other surfaces reasonably well. In the approach here, we keep increasing the cross section level until we find a section with a stable pattern. This process is shown in Figure~\ref{fig:crosscuts2} at the example of bullet 1-1 from barrel 3.
`Stability' is here defined as two aligned cross sections chosen 12.5$\mu m$ apart having a cross-correlation of at least 0.9.

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\linewidth}
\caption{Loess smooth at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{Using a rolling median peaks and grooves are identified for each cross section. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  lof <- processBullets(paths = images[c(5,7)], x = 100)
  lof <- bulletSmooth(lof)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

@@

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bullet~.) +
    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\linewidth}
    \caption{Rectangles in the back identify a line on one of the bullets.  Matching striation marks are indicated by color filled rectangles and arked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Cross section values (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between cross sections (c).}
\end{figure}
%\afterpage{\clearpage}

@@

Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet and pick the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet. 
%\hh{(assumption here: lower cross section levels show better expression of individual marks XXX checked XXX crosscuts-and-sd.R)}
